---
title: "Estratégia em Finanças - Artigo"
author: "A. Schmidt"
date: "May 5, 2019"
header-includes:
   - \usepackage{bigints}
   - \usepackage[brazil]{babel}
   - \usepackage{graphicx}
   - \usepackage{amsmath}
output: html_document
---

# Introduction

This code downloads and analyzes the data for my term paper for the "Estratégia em Finanças" course, taken in 2019-2. 

My to-do list is:

* Write an script to download the data from the [30 mining companies listed on NYSE](https://www.investing.com/stock-screener/?sp=country::5|sector::a|industry::16|equityType::a|exchange::1%3Ceq_market_cap;1);
    * *Stretch goal*: get some control data, for example prices of mining ore, SP500 index
* Clear data to get the daily prices;
    * *Stretch goal*: Get the list of the calendar business days to make easier to cross with the disaster;
    * *Stretch goal*: Compare the prices of the Vale do Rio Doce stocks in B3 (former BM&F) and NYSE;
* Find the cut point in the series to split the data for two separate sets: one for the Mariana disaster (05/11/2015) and another for Brumadinho disaster (25/01/2019);
* Get the package and documentation for the synthetic control;
    * Run some cluster analysis;
    * See if the differences (if exist) are equal between the two disasters.

## Downloading packages

If mirror 10 ("UFRJ") is offline, try changing `ind = 10` to `ind = 1`. Do the same thing if you get an error installing the `BETS` package.

```{r, warning = FALSE, message = FALSE}
chooseCRANmirror(graphics = FALSE, ind = 10)
if (!require("pacman")) install.packages("pacman")
pacman::p_load(ggplot2, forecast, BETS, seasonal, seasonalview, lubridate, zoo, stargazer, gridExtra, reshape2, ggfortify, RColorBrewer, scales, quantmod, PerformanceAnalytics, strucchange, knitr, grid, ggpubr, gdata, vars, urca, compiler, DescTools, kableExtra, readxl, tseries) 
```

# Downloading and preparing data

First, I need to get the symbols list, they are in a github repo.

```{r, message = FALSE, warning = FALSE}
#sheet   <- read.csv("D:\\Onedrive - Aisha\\OneDrive\\Documentos\\Graduação Economia UDESC\\2019-1\\Estratégia em Finanças - Daniel\\Artigo da disciplina\\Mining companies listed on NYSE.csv")
sheet <- read.csv("https://raw.githubusercontent.com/aishameriane/Vale-Stock-and-Disasters/master/Mining%20companies%20listed%20on%20NYSE.csv?token=AAVGJTU33E37R27QB7BSAJK43DKLO")
colnames(sheet)[1] <- "Name"
sheet <- sheet[-c(19,5),] # for some reson the thing does not understand the -P notation for the preferential stocks, so for now I removed them
symbols <- as.character(sheet[,2])
```

```{r, message = FALSE, warning = FALSE}
getSymbols(symbols, src = 'yahoo', from = "2015-01-01", 
             auto.assign = TRUE, warnings = FALSE) 
prices <- merge(BHP, BBL, RIO, VALE, SCCO, FCX, ACH, ARNC, VEDL, AA, CCJ, TRQ, CLF, CMC, HBM, HSC, NEXA, MTL, NXE, LAC, UEC, PLM, TGB, URG, PLG, UAMY, GMO, XPL)
prices <- prices[,grepl( "Close" , names(prices) )]
colnames(prices) <- as.character(sheet[,1])
returns <- na.omit(Return.calculate(prices, method = "log"))*100
```
